---
title: "Study Guide"
author: "Katie Volpert"
date: "2023-09-27"
output: pdf_document
---

VARIABLE TYPES:

Categorical — usually strings, identify which category something falls into 
  Ordinal — ordered categorical variables (ex. education levels)
  Nominal — unordered
  Binary — 0 or 1, yes or no, 2 options
Quantitative — numbers
  Continuous — the values in between the values are relevant
  Discrete — the in between values don't matter

QUANTITATIVE ANALYSIS:

```{r, include = TRUE}

# Initial analysis
summary() # tells you the min, max, mean, median, and quartiles (only for quantitative)
table() 
# one variable tells you the count of each value (mostly used for categorical)
# two variables creates a contingency table 
prop.table(table()) # tells you the proportion of each value (mostly used for categorical)

# na.rm = T to exclude NAs in the calculations

# Center
median()
mean()

# Shape
# Modes — number of humps in the histogram
  # Unimodal = one
  # Bimodal = two 
  # Multimodal = multiple
  # Uniform = no peaks
# Symmetry
# Outliers
# Skew
  # Skewed right when the tail is on the right
  # Skewed left when the tail is on the left

# Spread
range()
sd() # standard deviation (units [average of the squared differences] from the mean)
var() # variance (the average of all data points in a group)
IQR() # quartiles

# ask if there is further analysis needed to quantitatively analyze a variable

```

GRAPHS:

```{r, include = TRUE}

install.packages("ggplot2")
library(gglot2) # always load the ggplot 2 package

# Scatterplot
ggplot(data = data, aes(x = x, y = y, color = z)) +
  geom_point()

# color is the third variable

# Barplot
ggplot(data = data, aes(x = x, y = y, fill = z)) +
  geom_bar(stat = "identity")

# Histogram
ggplot(data = data, aes(x = x)) +
  geom_histogram(position = "identity")

# Boxplot
ggplot(data = data, aes(x = x)) +
  geom_boxplot()

# Graph + Modifications
ggplot(data = data, aes(x = x, y = y, fill = z)) +
  geom_point() +
  ggtitle("Title") +
  xlab("X Axis Name") +
  ylab("Y Axis Name") + 
  theme_classic()

```

CORRELATIONS:

```{r, include = TRUE}

# Correlation
cor(x = x, y = y, method = "pearson")

# Correlation strengths guide
# 0.9-1.00: Very strong
# 0.7-0.9: High correlation
# 0.5-0.7: Moderate correlation
# 0.3-0.5: Low correlation
# 0.0-0.3: Negligible correlation

```

